### 2017-11-20 manuscript.rmd
spelled out OC

While empirical support for the ortholog conjecture has been mixed, theoretical work has suggested several mechanisms that could lead to increased rates of evolutionary change in expression and other gene traits following gene duplication. This has been a primary motivation for the extensive discussion of the OC. In his seminal work, Ohno [@Ohno:1970wm] outlined three outcomes of gene duplication, now often referred to as neofunctionalization (one gene copy can take on new functions while the other retains the old functions), subfunctionalization (each gene copy has a subset of the functions present in their most recent common ancestor), and conservation (each gene copy retains the same function). Neofunctionalization was the primary focus of early work on gene duplication, though it was not found to be a widespread outcome of duplication [@Hughes:1994ec]. The highly influential duplication-degeneration-complementation (DDC) model [@Force:1999tg; @Lynch:2000um] suggested that subfunctionalization could be a better explanation for the retention of both copies of a gene. It is built on the idea that gene copies that have each lost different functions would both be needed to fulfill all the functions present in their most recent common ancestor. Tests of the DDC model have had mixed results [@Huminiecki:2004dc]. A growing body of work suggests that retention and differentiation of genes is due to a variety of different mechanisms in different contexts, and that we should not necessarily expect one process to dominate over others [@Hahn:2009by]. Our lack of support for the ortholog conjecture here suggests that whatever mechanisms lead to the retention of duplicated genes in the organisms considered here does not lead to changes in the tissue-specificity of gene expression as summarized by tau.

The DDC model is particularly interesting to further consider in this context since the summary statistic tau describes tissue specificity of expression. In the context of tissue-specific expression, the OC predicts a greater rate of evolutionary change in tau following duplication, without specifying the direction of the change. The DDC model predicts a specific type of change in expression following duplication -- subfunctionalization, which would increase tissue specificity, and therefore tau, following duplication. To test for this more specific predictions of DDC, we conducted additional analyses to see if there is a greater rate of increase in tau following duplication. For each branch in each gene tree, we inferred the change in tau and standardized this change by branch lengths. We found no evidence of increased tau following duplication relative to speciation (Wilcoxon $P=`r  edge_p %>% round( 3 )`$). This result also holds when we discarded the `r k_percentile` of genes with the lowest phylogenetic signal (Wilcoxon $P=`r  edge_pass_k_p %>% round( 3 )`$). These results are consistent with another recent study that found little support for neofunctionalization or subfunctionalization in expression of duplicated human genes, and instead proposed that dosage compensation drives most changes in expression following duplication in mammals [@Lan:2016ix]. Like the OC, DDC is not an all-or-nothing hypothesis. We fail to find support for it in these tissues in these species with the tau expression summary statistic, but it may hold in other contexts. DDC could still play a role in the retention of these genes, but with respect to other attributes such as biochemical functions rather than tissue-specific expression.

>>>>

While empirical support for the ortholog conjecture has been mixed, theoretical work has suggested several mechanisms that could lead to increased rates of evolutionary change in expression and other gene traits following gene duplication. This has been a primary motivation for the extensive discussion of the ortholog conjecture. In his seminal work, Ohno [@Ohno:1970wm] outlined three outcomes of gene duplication, now often referred to as neofunctionalization (one gene copy can take on new functions while the other retains the old functions), subfunctionalization (each gene copy has a subset of the functions present in their most recent common ancestor), and conservation (each gene copy retains the same function). Neofunctionalization was the primary focus of early work on gene duplication, though it was not found to be a widespread outcome of duplication [@Hughes:1994ec]. The highly influential duplication-degeneration-complementation (DDC) model [@Force:1999tg; @Lynch:2000um] suggested that subfunctionalization could be a better explanation for the retention of both copies of a gene. It is built on the idea that gene copies that have each lost different functions would both be needed to fulfill all the functions present in their most recent common ancestor. Tests of the DDC model have had mixed results [@Huminiecki:2004dc]. A growing body of work suggests that retention and differentiation of genes is due to a variety of different mechanisms in different contexts, and that we should not necessarily expect one process to dominate over others [@Hahn:2009by]. Our lack of support for the ortholog conjecture here suggests that whatever mechanisms lead to the retention of duplicated genes in the organisms considered here does not lead to changes in the tissue-specificity of gene expression as summarized by tau.

The DDC model is particularly interesting to further consider in this context since the summary statistic tau describes tissue specificity of expression. In the context of tissue-specific expression, the ortholog conjecture predicts a greater rate of evolutionary change in tau following duplication, without specifying the direction of the change. The DDC model predicts a specific type of change in expression following duplication -- subfunctionalization, which would increase tissue specificity, and therefore tau, following duplication. To test for this more specific predictions of DDC, we conducted additional analyses to see if there is a greater rate of increase in tau following duplication. For each branch in each gene tree, we inferred the change in tau and standardized this change by branch lengths. We found no evidence of increased tau following duplication relative to speciation (Wilcoxon $P=`r  edge_p %>% round( 3 )`$). This result also holds when we discarded the `r k_percentile` of genes with the lowest phylogenetic signal (Wilcoxon $P=`r  edge_pass_k_p %>% round( 3 )`$). These results are consistent with another recent study that found little support for neofunctionalization or subfunctionalization in expression of duplicated human genes, and instead proposed that dosage compensation drives most changes in expression following duplication in mammals [@Lan:2016ix]. Like the ortholog conjecture, DDC is not an all-or-nothing hypothesis. We fail to find support for it in these tissues in these species with the tau expression summary statistic, but it may hold in other contexts. DDC could still play a role in the retention of these genes, but with respect to other attributes such as biochemical functions rather than tissue-specific expression.
